A dataset of Scratch programs,\\Scraped, Shaped and Scored

### 2017-02-17 MSR-Scratch.tex
Added limitation
Add Dr. Scratch text and a couple of comments

\newcommand{\fenia}[1]{\emph{\color{blue}Fenia says: #1}}
\newcommand{\felienne}[1]{\emph{\color{green}Felienne says: #1}}
\newcommand{\grex}[1]{\emph{\color{yellow}Gregorio says: #1}}
\newcommand{\jemole}[1]{\emph{\color{red}Jesús says: #1}}

@@

A number of works in the computing education research field attempt to assess the programming skills that novice programmers develop in the Scratch environment.
Some utilize program data collected during specific programming courses (e.g., \cite{meerbaum-salant_learning_2010, wilson_evaluation_2012, Maloney_2008}), while others utilize the dataset made available by the Lifelong Kindergarten Group at the MIT Media Lab, which contains data for Scratch projects created until 2012 (e.g., \cite{fields_2014, yang_2015, Dasgupta_2016}).
In addition to identifying indications of learning of programming concepts, static analysis of Scratch programs has also been performed for identifying code smells and bad programming practices within Scratch programs \cite{Meerbaum_habits_2011, Aivaloglou_2016}, and automated quality assessment tools have been proposed (e.g. Hairball \cite{boe_hairball:_2013} and Dr. Scratch \cite{moreno_automatic_2014}).

@@

\subsection{Calculation of programming mastery scores}
\fenia{Description of how the mastery metrics for       
	[Abstraction]
	,[Parallelism]
	,[Logic]
	,[Synchronization]
	,[FlowControl]
	,[UserInteractivity]
	,[DataRepresentation]
	,[Mastery]
	,[Clones]
	,[CustomBlocks]
	,[InstancesSprites]
	were calculated}

@@

\begin{table}[]
	\centering
	\begin{tabular}{llp{5.4cm}}
		\textbf{Table}& \textbf{Key} & \textbf{Attribute(Description)}\\
		\hline
		\textbf{\texttt{project}} & PK & \texttt{p\_id} (Scratch project ID)\\
		&  & 							\texttt{project-name} (name given to project)\\
		&  & 							\texttt{username} (author's Scratch 
		username)\\
		&  & 							\texttt{total-views} (project views number)\\
		&  & 							\texttt{total-remixes} (project remixes number)\\
		&  & 							\texttt{total-favorites} (total users favoriting)\\
		&  & 							\texttt{total-loves} (total users `loving' the project)\\
		& & 							\texttt{is-remix} (calculated column, true if project is a remix of another one)\\
		\hline
		\textbf{\texttt{script}} & PK & \texttt{script\_id} (auto increment)\\
		& FK & 							\texttt{project\_id} (\texttt{\scriptsize{*\ldots1 project:p\_id}})\\
		& & 							\texttt{sprite-type} (one of [sprite, stage, procDef])\\
		& & 							\texttt{sprite-name} (name sprite the script is in)\\
		& & 							\texttt{script-rank} (project-level ranking of script)\\
		& & 							\texttt{coordinates} (x-y location of script in Scratch editor)\\
		& & 							\texttt{total-blocks} (calculated column, number of blocks comprising the script)\\
		\hline
		\textbf{\texttt{procedure}} & PK,FK & \texttt{script\_id} (\texttt{\scriptsize{1\ldots1 script:script\_id}})\\
		& & 							\texttt{proc-name} (name given to custom block)\\
		& & 							\texttt{total-args} (number of input arguments)\\
		\hline
		\textbf{\texttt{block}} & PK,FK  & \texttt{script\_id} (\texttt{\scriptsize{*\ldots1 script:script\_id}})\\
		& PK & \texttt{block-rank} (script-level ranking of block)\\
		& FK & \texttt{block-type} (\texttt{\scriptsize{*\ldots1 blockType:b-type}})\\
		&  & \texttt{parameter-1} (value of 1st input parameter)\\
		&  &\vdots\\
		&  & \texttt{parameter-24} (value of 24th input parameter)\\
		\hline
		\textbf{\texttt{blockType}} & PK & \texttt{b-type} (Scratch name of predefined block)\\
		&  & \texttt{category} (Scratch block category)\\
		&  & \texttt{shape} (One of the 8 block shapes in the Scratch interface)\\
		&  & \texttt{is-input} (true if the block receives user input)\\
		\hline
		\textbf{\texttt{remix}} & PK  & \texttt{remix\_id} (\texttt{\scriptsize{project\_id}} of created project)\\
		&  & \texttt{from-project\_id} (\texttt{\scriptsize{project\_id}} of original project)\\
		\hline
		\textbf{\texttt{grades}} & PK,FK & \texttt{project\_id} (\texttt{\scriptsize{1\ldots1 project:p\_id}})\\
		& & \texttt{Abstraction} (\fenia{short descriptions})\\
		& & \texttt{Parallelism} ()\\
		& & \texttt{Logic} ()\\
		& & \texttt{Synchronization} ()\\
		& & \texttt{FlowControl} ()\\
		& & \texttt{UserInteractivity} ()\\
		& & \texttt{DataRepresentation} ()\\
		& & \texttt{Mastery} ()\\
		& & \texttt{Clones} ()\\
		& & \texttt{CustomBlocks} ()\\
		& & \texttt{InstancesSprites} ()\\
		\hline
	\end{tabular}
	\caption{Database schema: Tables and Attributes}
	\label{tbl-dbschema}
\end{table}

@@

\fenia{Should we add some stats about the Dr Scratch master scores?}

@@

Another limitation of this dataset is that we did not scrape a random sample of Scratch projects, but the 250,000 most recent ones during the time that we run the scraper.
It could be the case that the programming habits of Scratch users are changing over time.
However, we counterbalanced that by collecting a large dataset which comprises around 1.3\% of all 19 million shared Scratch projects, and is still the most recent one to be made available to the research community.

>>>>

\newcommand{\fenia}[1]{\emph{\color{blue}Fenia says: #1}}
\newcommand{\felienne}[1]{\emph{\color{green}Felienne says: #1}}
\newcommand{\grex}[1]{\emph{\color{yellow}Gregorio says: #1}}
\newcommand{\jemole}[1]{\emph{\color{red}Jes\'us says: #1}}

@@

A number of works in the computing education research field attempt to assess the programming skills that novice programmers develop in the Scratch environment.
Some utilize program data collected during specific programming courses (e.g., \cite{meerbaum-salant_learning_2010, wilson_evaluation_2012, Maloney_2008}), while others utilize the dataset made available by the Lifelong Kindergarten Group at the MIT Media Lab \jemole{perhaps we could cite the dataset itself~\cite{2017_scratch_dataset}}, which contains data for Scratch projects created until 2012 (e.g., \cite{fields_2014, yang_2015, Dasgupta_2016}).
In addition to identifying indications of learning of programming concepts, static analysis of Scratch programs has also been performed for identifying code smells and bad programming practices within Scratch programs \cite{Meerbaum_habits_2011, Aivaloglou_2016}, and automated quality assessment tools have been proposed (e.g. Hairball \cite{boe_hairball:_2013} and Dr. Scratch \cite{moreno_automatic_2014}).

@@

\subsection{Calculation of programming mastery scores}
%\fenia{Description of how the mastery metrics for       
%	[Abstraction]
%	,[Parallelism]
%	,[Logic]
%	,[Synchronization]
%	,[FlowControl]
%	,[UserInteractivity]
%	,[DataRepresentation]
%	,[Mastery]
%	,[Clones]
%	,[CustomBlocks]
%	,[InstancesSprites]
%	were calculated}
The projects were analyzed with Dr. Scratch, a tool that statically inspects Scratch projects source code to assign scores on seven dimensions of computational thinking: abstraction and problem decomposition, logical thinking, synchronization, parallelism, algorithmic notions of flow control, user interactivity and data representation. These dimensions are given a value from 0 to 3. The mastery score is the aggregated punctuation of these dimensions, and therefore ranges from 0 to 21.

In addition, Dr. Scratch also detects intra-project software cloning, i.e. repetition of code; the use of custom blocks, which is the way functionality can be reused in Scratch programs; and the use of instances of sprites, a feature labeled in Scratch as clone creation. In consequence, these values are also included in the dataset.

@@

\begin{table}[]
	\centering
	\begin{tabular}{llp{5.4cm}}
		\textbf{Table}& \textbf{Key} & \textbf{Attribute(Description)}\\
		\hline
		\textbf{\texttt{project}} & PK & \texttt{p\_id} (Scratch project ID)\\
		&  & 							\texttt{project-name} (name given to project)\\
		&  & 							\texttt{username} (author's Scratch 
		username)\\
		&  & 							\texttt{total-views} (project views number)\\
		&  & 							\texttt{total-remixes} (project remixes number)\\
		&  & 							\texttt{total-favorites} (total users favoriting)\\
		&  & 							\texttt{total-loves} (total users `loving' the project)\\
		& & 							\texttt{is-remix} (calculated column, true if project is a remix of another one)\\
		\hline
		\textbf{\texttt{script}} & PK & \texttt{script\_id} (auto increment)\\
		& FK & 							\texttt{project\_id} (\texttt{\scriptsize{*\ldots1 project:p\_id}})\\
		& & 							\texttt{sprite-type} (one of [sprite, stage, procDef])\\
		& & 							\texttt{sprite-name} (name sprite the script is in)\\
		& & 							\texttt{script-rank} (project-level ranking of script)\\
		& & 							\texttt{coordinates} (x-y location of script in Scratch editor)\\
		& & 							\texttt{total-blocks} (calculated column, number of blocks comprising the script)\\
		\hline
		\textbf{\texttt{procedure}} & PK,FK & \texttt{script\_id} (\texttt{\scriptsize{1\ldots1 script:script\_id}})\\
		& & 							\texttt{proc-name} (name given to custom block)\\
		& & 							\texttt{total-args} (number of input arguments)\\
		\hline
		\textbf{\texttt{block}} & PK,FK  & \texttt{script\_id} (\texttt{\scriptsize{*\ldots1 script:script\_id}})\\
		& PK & \texttt{block-rank} (script-level ranking of block)\\
		& FK & \texttt{block-type} (\texttt{\scriptsize{*\ldots1 blockType:b-type}})\\
		&  & \texttt{parameter-1} (value of 1st input parameter)\\
		&  &\vdots\\
		&  & \texttt{parameter-24} (value of 24th input parameter)\\
		\hline
		\textbf{\texttt{blockType}} & PK & \texttt{b-type} (Scratch name of predefined block)\\
		&  & \texttt{category} (Scratch block category)\\
		&  & \texttt{shape} (One of the 8 block shapes in the Scratch interface)\\
		&  & \texttt{is-input} (true if the block receives user input)\\
		\hline
		\textbf{\texttt{remix}} & PK  & \texttt{remix\_id} (\texttt{\scriptsize{project\_id}} of created project)\\
		&  & \texttt{from-project\_id} (\texttt{\scriptsize{project\_id}} of original project)\\
		\hline
		\textbf{\texttt{grades}} & PK,FK & \texttt{project\_id} (\texttt{\scriptsize{1\ldots1 project:p\_id}})\\
		& & \texttt{Abstraction} (Dr. Scratch score for abstraction and problem decomposition)\\
		& & \texttt{Parallelism} (Dr. Scratch score for parallelism)\\
		& & \texttt{Logic} (Dr. Scratch score for logical thinking)\\
		& & \texttt{Synchronization} (Dr. Scratch score for synchronization)\\
		& & \texttt{FlowControl} (Dr. Scratch score for algorithmic notions of flow control)\\
		& & \texttt{UserInteractivity} (Dr. Scratch score for user interactivity)\\
		& & \texttt{DataRepresentation} (Dr. Scratch score for data representation)\\
		& & \texttt{Mastery} (Dr. Scratch total mastery score)\\
		& & \texttt{Clones} (Number of software clones)\\
		& & \texttt{CustomBlocks} (Number of user defined blocks)\\
		& & \texttt{InstancesSprites} (True if project makes use of Scratch clones)\\
		\hline
	\end{tabular}
	\caption{Database schema: Tables and Attributes}
	\label{tbl-dbschema}
\end{table}

@@

\fenia{Should we add some stats about the Dr Scratch master scores?}
\jemole{I don't know if we have space for this, but here it goes a proposal: As shown in Table~\ref{tbl-size}, not all the projects were successfully analyzed with Dr. Scratch. This is due to problems with non-official extension blocks and with some non-ascii characters in sprites names. Of those correctly assessed projects, as depicted in Table~\ref{tbl-stats}, the mastery score of the median project is 8, which is considered a medium level of development of computational thinking skills.}

@@

Another limitation of this dataset is that we did not scrape a random sample of Scratch projects, but the 250,000 most recent ones during the time that we run the scraper.
It could be the case that the programming habits of Scratch users are changing over time.
However, we counterbalanced that by collecting a large dataset which comprises around 1.3\% of all 19 million shared Scratch projects, and is still the most recent one to be made available to the research community.
It must also be noted that the projects in the dataset are the ones that their authors shared publicly when the scraper run. The dataset contains cases of projects that are no longer publicly shared and are thus no longer accessible through the Scratch web interface.
